---
title: Test if I can knit PDF
author: William Komu
document: PDF-doc
---

## Loading the Requisite Packages

Loading packages is crucial to avoid gettting into issues when you start your analysis. Thus, I always start a new doc with a chuck importing all the packages that I would use!

```{r}
library(tidyverse)
```

Sure Enough, this one is knitting perfectly! Thus, the one that is not knitting must be labouring with some weird settings!!!

## Add More Details

1. ADD THE SETUP CHUCK

```{r setup, include=TRUE}
knitr::opts_chunk$set(
    echo = FALSE,
    message = FALSE,
    warning = FALSE
)
# This setup covers the whole document!
# Forbids the code from displaying!

```

Let's try and knit with the added setup chuck above!

**It still knits**

2. ADD THE emojis

<!-- `r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")` -->

Let's try and knit with emojis:

The emojis have killed it! Let me see if I can ***settle this issue out***!
- First, comment the emojis out!

It prints with the emojis commented!

- add emojis as standalone code chucks

```{r}

emo::ji("star")

```

# TENTATIVE FINDINGS

The emojis are preventing me from knitting to pdf! 
