---
title: "Global Tempature Anomaly"
subtitle: "Replicating NASA's Graph"
author: 'William Komu'
date: '`r Sys.Date()`'
doc: pdf_doc
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
        echo = FALSE,
        message = FALSE,
        warning = FALSE
)
```


```{r}
library(tidyverse)
library(janitor)
library(emo)
library(scales)
library(glue)
```

This exercise is inspired by Pat Schloss of the [Riffomonas Project](https://riffomonas.org/code_club/2022-05-19-temperature-index). It is an attempt to replicate the NASA's g

## Import data and tidy it

```{r import and tidy data, echo = TRUE}
t_diff <- read_csv("/home/iamlnx/Documents/GitHub/nasoma_R/climate/data/GLB.Ts+dSST.csv", skip = 1, na = "***") %>%
        select(Year, "J-D") %>%
        rename(t_diff = "J-D") %>%
        clean_names()
```
## Plot

This plot mimics NASA's [GLOBAL LAND-OCEAN TEMPERATURE INDEX](https://climate.nasa.gov/vital-signs/global-temperature/). The graph indicates the deviation of the global avg temperature from a normalised mean!

```{r plot code chuck, echo = TRUE}
t_diff %>%
        ggplot(aes(year, t_diff)) +
        geom_line(aes(color = "1"), size = 0.5, show.legend = FALSE) +
        geom_point(fill = "white", aes(color = "1"), shape = 21, show.legend = TRUE) +
        geom_smooth(aes(color = "2"), span = .15, size = 0.5, se = FALSE, show.legend = FALSE) +
        theme_light() +
        scale_x_continuous(breaks = seq(1880, 2020, 20), expand = c(0, 0)) +
        scale_y_continuous(limits = c(-0.5, 1.5), expand = c(0, 0)) +
        scale_color_manual(
                name = NULL,
                breaks = c(1, 2),
                values = c("gray", "black"),
                labels = c("Annual mean", "Loess smoothing"),
                guide = guide_legend(override.aes = list(shape = 15, size = 5))
        ) +
        labs(
                title = "GLOBAL LAND-OCEAN TEMPERATURE INDEX",
                x = "YEAR",
                y = "Temperature Anomaly (C)",
                subtitle = "Data source: NASA's Goddard Institute for Space Studies (GISS).\nCredit: NASA/GISS"
        ) +
        theme(
                axis.ticks = element_blank(),
                plot.title.position = "plot",
                plot.title = element_text(margin = margin(b = 10), color = "red", face = "bold"),
                plot.subtitle = element_text(size = 8, margin = margin(b = 10)),
                legend.position = c(0.15, 0.9),
                legend.title = element_text(size = 0),
                legend.key.height = unit(10, "pt"),
                legend.margin = margin(0, 0, 0)
        )
```

       <!-- `r emo::ji("fire")``r emo::ji("fire")` `r emo::ji("fire")` #EarthOnFire `r emo::ji("fire")` `r emo::ji("fire")``r emo::ji("fire")` -->

# Interpretation of the Code Chuck
## Scale the x and y-axis
### x-axis
I was looking into this the other day. All my plots have bothered me... for not capturing all the x values, which means some of the latter years are ignored! To fix, let's use:

        `scale_x_continuous(breaks = seq(1880, 2020, 20), expand = c(0, 0))`

The first part of the code, `breaks = seq(1880, 2020, 20)`, formats the x-values starting from **1880-2020**, with 20-year intervals, and forces the x-values to start and stop at the plot area corners - ***no padding***!

The other part `expand = c(0, 0)` instructs R to get rid of spaces between the plot's edges!

### y-axis

The layer `scale_y_continuous(limits = c(-0.5, 1.5))`sets the `ymin=-0.5` and `ymax=1.5`! I will look to using the ymin and ymax syntax later! I think it is used. By the scale_y_continuous is much better for me coz I have the one for colors mastered.

### Showcasing the depth of the `theme()` function

The `theme()` code will allow us to

- Remove the ticks!

        axis.ticks = element_blank()

- Align tittle left extreme edge:

        plot.title.position = "plot"

***Without this setup, the title will align with the left border of the chart area!**

- Add space between title and subtitle, color title red and bold it:

        plot.title = element_text(margin = margin(b = 10), color = "red", face = "bold")

- Size the subtitle, create a buffer space between it it and top of chart area

        plot.subtitle = element_text(size = 8, margin = margin(b = 10))

- Place legend on the top left side of the chart area

        legend.position = c(0.15, 0.9)

- A Trick for something! HAHA

        legend.title = element_text(size = 0)

- Fix legend space between legend keys

        legend.key.height = unit(10, "pt")

- Refine legend postion relative to the left, top edges

# Save PNG defaults



```{r save plot, echo=TRUE}
ggsave("global_temparature_index.png")
```

# TEMPERATURE BAR PLOT
### Plotting the global temperature index as bars using ggplot2 and NASA GISS data (CC215)

```{r}
t_data <- t_diff
```
```{r}
annotation <- t_data %>%
        slice(1, n()) %>%
        mutate(t_diff = 0, x = year + c(-5, 5))

annotation

max_diff <- format(round(max(t_data & t_diff), 1), nsmall = 1)
max_diff
```

```{r}
t_data %>% ggplot(aes(year, t_diff, fill = t_diff)) +
        geom_col(show.legend = FALSE) +
        geom_text(data = annotation, aes(x = x, label = year), color = "white") +
        geom_text(
                x = 1880, y = 1,
                label = glue("Global temparatures have increased by over {max_diff}\u00B0C since 1880"),
                hjust = 0,
                color = "white"
        ) +
        # scale_fill_gradient2(low = "darkblue", mid = "white", high = "darkred", midpoint = 0, limits = c(-0.5, 1.5)) +
        scale_fill_stepsn(
                colors = c("darkblue", "white", "darkred"),
                values = rescale(c(min(t_data$t_diff), 0, max(t_data$t_diff))),
                limits = c(min(t_data$t_diff), max(t_data$t_diff)),
                n.breaks = 7
        ) +
        theme_void() +
        theme(
                plot.background = element_rect(fill = "black"),
                legend.text = element_text(color = "white")
        )

ggsave("temparature_bar_plot.png")
```


# Genious annotation method

I gotta give it to you Professor. I have just told a friend you must have been a core member of those that created some of these function! The slice, mutate, and mutate again are genious forms of annotation!
