---
title: Dealing with Notorious Data Sets
subtitle:       An ntroduction to thinking like a data analyst
author: William Komu
date: '`r Sys.Date()`'
document: pdf_doc
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE,
    message = FALSE,
    warning = FALSE
)
# This setup covers the whole document!
# Forbids the code from displaying!

```

***I have noticed that the code chuck above setup the entire document such that we don't have to keep repeating the `echo=FALSE` line at every insertion of a code chuck***

***Oooooo...f course you won't see it! So, lemme give you a test:***

## The Knitr Setup code

    {r setup, include=FALSE}
    knitr::opts_chunk$set(
        echo = FALSE,
        message = FALSE,
        warning = FALSE
    )
    # This setup covers the whole document
    # We cannot display code

`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`


**This is an emoji bomb!!!!! The inclusion of this bomb is just a showoff that carries no meaningful meaning. It seems this is meanigless avenue!!!**

`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`

Of course, I can do other emojis!

`r emo::ji("monkey")`
`r emo::ji("monkey")`
`r emo::ji("monkey")`
`r emo::ji("poop")`
`r emo::ji("poop")`
`r emo::ji("poop")`
`r emo::ji("poop")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("star")`
`r emo::ji("star")`
`r emo::ji("star")`

Apparently, repeating the face consequentively generates various faces

`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`

## The Case Study

This exercise offers us an opppotunity to apply a series of techniques that we have acquired over the course of our code club sessions!

**The Knitr Setup Chuck** above contains `echo = FALSE`, which hides the code that generates a given outcome in the knitted document! If you want the code to appear, you could set it per chuck by substituting the `echo = FALSE` wth `echo = TRUE`!


Oooh! My bad!! You can't see it coz it has echoed itself off!

# Pseudocode

Let us create an outline of what this exercise entails:
1. Read the csv and excel filters
2. Subset the only graduates from MIU
3. Subset the portion of the excel that indicates the code for degrees aarded!
4. filter
5. rename
6. clean
7. JOIN
   1. Perharps the most crucial part of the exercise!
   2. simple if you know what you want and how to get it!


### Load or install packages

```{r}
library(tidyverse)
library(readxl)
library(emo)
library(janitor)
library(knitr)
library(rmarkdown)

# install.packages("janitor")
# remotes::install_github("hadley/emo")

```


Let's test emojis in `.rmd`:

```{r, echo=FALSE}
emo::ji("star")

emo::ji("poop")
```

Given the setup code above has `echo=FALSE` for the whole document, we can only report the code chuck if we override the setup instructions as follows:

```{r, echo=TRUE}
emo::ji("star")

emo::ji("poop")
```

## Import the Data
### The Entire Dataset

```{r imports entire degree issued dataset}

degrees <- read_csv("/home/iamlnx/Documents/nasomaR/data/c2018_a.csv")

```

### The Data Dictionary

Let's now load the other data set, whihc is a dictionary of explaining the data sets. This one is in .xlsx form. Thus, we are going to use `read_excel` function. This function lives in the `readxl` universe! Or does it? This is a topic that we will examine in a few, eos or seconds!

The chuck below calls the data set and filter only the part of the data that we need - that is, the one that explains the above codes for degree!

```{r}
data_dict <- read_excel("/home/iamlnx/Documents/nasomaR/data/c2018_a.xlsx", sheet = "Frequencies")

```

### Printing the Two Tables - `degrees` and `degree_name`
```{r}
knitr::kable(degrees[1:10, 1:10], caption = "A subset of degrees in the data set")
```

```{r}
knitr::kable(data_dict[1:8, 1:6], caption = "A subset of degrees in the data set")
```

The data is shitty, huh?

### Filter only MIU CIPCODE=170976

The code below filters the Michichan University (MIU), then arrange (which is redundant because it offers no addtional value and it is just there to help us ensure there is a value `99`), then filters out the CIPCODE 99 since it is just a grand sum of some shiet I don't know, then filters only the first class degrees that the MIU issued in 2018!

Then return the count of by degree_type

```{r}
first_c <- degrees %>%
    filter(UNITID == 170976) %>%
    arrange(desc(CIPCODE)) %>%
    filter(CIPCODE != 99 & MAJORNUM == 1) %>%
    count(AWLEVEL) %>%
    rename("degree_code" = AWLEVEL)
```

If I print the table, we can see it is a shitty one

### Filter
```{r}
degrees_names <- data_dict  %>%
select(codevalue, valuelabel) %>%
    rename(
        "degree_code" = codevalue,
        "degree_name" = valuelabel
    ) %>%
    filter(degree_code %in% c("5", "6", "7", "8", "17", "18"))
```

The chuck below filters the data by the MIU's `CIPCODE`, reporting first class honors awards. Then filters out cipcode +, which is a grand total!

### Bad Robot

Unfortunately, the data degree are coded. One has to correspond the codes with the other set of data to `` to reposrt the degree names!

```{r}
knitr::kable(first_c, caption = "The count of first class honors that MIU issued by degree_code"
)
```

```{r}
knitr::kable(degrees_names, caption = "The the degrees_code and respective degrees_names that MIU issued"
)
```

## JOINING THE DATA FRAMES

### Menial Work

As you can note we have two dfs now. This is not helpful, is it? Definately, not! Thus, we need a way to combine the two to make this stuff meaningful even for those that have no idea what the dfs entail.

This is a rather mundane process, but it is not as simple for a layman as it is for an R master, that I claim, as my new **moniker**.

I mean, you could just get the degrees awarded and correspond them with the values, manually! That is, recode the values one by one! But that is a lot of work. Let's use the power of R to do it!

But aint that tedious??? It really is!!

### Demonstrate the Power of R to Deal with Untidyness

```{r, echo=FALSE}

final_miu_first <- data.frame(degrees_names, c(first_c)) %>%
    rename("number_of_graduates"=n)

```

### Print Final Tables


```{r}
final_miu_first
```

This table has joined the set and we can tidy it by selecting only the names and the count:

```{r}
final_miu_first %>%
    select(degree_name, number_of_graduates)

```




# THE GUINEA PIG
## A Series of Face emojis

`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`
`r emo::ji("face")`


/home/iamlnx/Documents/GitHub/nasoma_R/code/generalR_s11.Rmd


<!-- Git Commands
git add .gitignore code/test.Rmd code/generalR_s11.Rmd

git commit -m "Initial commit of a .gitignire and the generalR files"
 -->
