---
title: 'Privy Practice: Visualization and Statistical Analysis of Global Life Expectancy World Bank Data in R'
author: 'William Komu'
date: '`r Sys.Date()`'
document:
  pdf_document:
    self_contained: false
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE,
    message = FALSE,
    warning = FALSE
)
options(browser = "firefox")
```

# Introduction

This exercise attempts to visualise the global life expectancy and conduct a variety of statistical tests to compare the life expectancy of among a select countries.
The primary purpose is to compare how life expectancy in Kenya has changes over the years!!

## Loading packages

Like any other R-related exercise, it is a prerequisite that we load the various packages that are going to enable the functionality that we need to successfully conduct this exercise!
Like usual, I will load the <font color="dodgerblue">`tidyverse`</font>.

## The environment

Like usual, I have almost all the required packages install on my <font color="dodgerblue">`nasoma_r` </font> env, which is a <font color="dodgerblue"> conda/mamba </font> environment!
### Loading packages The chuck below loads the packages that I need or could need as I take on this exercise!

```{r load_pkgs}
library(tidyverse)
library(janitor)
library(ggthemes)
library(lubridate)
library(gganimate)

```

## Color pallete

I had copied two color palette, which supposedly show empathy to those with color blindness, from [this site](example.com)!
They are based on a few colors with a base of `gray` and `black`!
I went on and saved them as snippets.

I load them below:

```{r load_my_color_palette}
cb_palette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
cbb_palette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

## Load Data

The data is a csv file that I downloaded from the WB data repo.

```{r loads_data}
my_gapminder <- read_csv("/home/iamlnx/Documents/GitHub/nasoma_R/global_life_exp/data/LifeExp_WorldBank.csv", skip = 4) %>%
    select(-"Country Code", -"Indicator Name", -"Indicator Code") %>%
    pivot_longer(-"Country Name", names_to = "year", values_to = "life_exp") %>%
    clean_names() %>%
    mutate(
        country = country_name,
        year = as.integer(year)
    ) %>%
    group_by(country) %>%
    select(country, year, life_exp) %>%
    drop_na()
```

# VISUALIZATION

## Life expectancy - Kenya alone

This one is not a refined copy.
I have hidden the code that generates it!

```{r}
ke_p <- my_gapminder %>%
    filter(
        country %in% c("Kenya")
    ) %>%
    group_by(country) %>%
    ggplot(aes(year, life_exp)) +
    geom_line(show.legend = FALSE) +
    geom_point(fill = "white", shape = 18, size = 2, show.legend = FALSE) +
    # shape enables the white fill!
    scale_x_continuous(breaks = seq(1960, 2030, by = 10), expand = c(0, 1)) +
    coord_cartesian(clip = "off") +
    scale_y_continuous(limits = c(45, 70), expand = c(0, 0)) +
    labs(
        title = "Kenya's life expectancy from 1960 to 2020",
        x = "Year",
        y = "Average life expectancy (years)",
        caption = "Data source: World Bank"
    ) +
    theme_clean() +
    theme(
        plot.title = element_text(margin = margin(b = 10)),
        plot.title.position = "plot",
        axis.text = element_text(margin = margin(b = 10))
    ) +
    transition_reveal(along = year)

ggsave("figures/kenya_life_exp_1960-2020.png", width = 6.5, height = 3.5, unit = "in")

# Animate and save .gif

animate(ke_p, width = 5.5, height = 4.5, unit = "in", res = 300)

anim_save("figures/kenya_lifeExp_line_animated_world_bank.gif")
```

## Comparison

### East Africa countries

```{r echo = TRUE}
my_gapminder %>%
    filter(country %in% c("Kenya", "Tanzania", "Uganda")) %>%
    drop_na() %>%
    ggplot(aes(year, life_exp, group = country)) +
    geom_line(aes(color = country), show.legend = TRUE, size = 0.5) +
    geom_point(aes(color = country), show.legend = TRUE, size = 1, shape = 16) +
    # geom_smooth(se = FALSE, span = 0.1) +
    scale_x_continuous(breaks = seq(1960, 2030, by = 10), expand = c(0, 1)) +
    coord_cartesian(clip = "off") +
    scale_y_continuous(limits = c(40, 70), expand = c(0, 0)) +
    scale_color_manual(breaks = c("Kenya", "Tanzania", "Uganda"), values = c(cbb_palette)) +
    labs(
        title = "Life expectancy from 1960 to 2020",
        subtitle = "East Africa",
        x = "Year",
        y = "Average life expectancy (in years)",
        caption = "Data source: World Bank"
    ) +
    theme_classic() +
    theme(
        plot.title = element_text(margin = margin(b = 10)),
        plot.title.position = "plot",
        axis.text = element_text(margin = margin(b = 10)),
        plot.subtitle = element_text(size = 10, margin = margin(b = 15)),
        legend.position = c(0.9, 0.15), # Moves the legend to the top right inside chart area!
        legend.title = element_blank() # Removes the legend title, country
    )

# Save `.png` of the plot

ggsave("figures/cbb_east_africa_life_exp_1960-2020.png", width = 5.5, height = 4.5, unit = "in")
```

## cb_palette - gray

The code below replicates the above chart, albeit using the <font color="gray">gray</font> variation color palette

```{r echo = TRUE}
my_gapminder %>%
    filter(country %in% c("Kenya", "Tanzania", "Uganda")) %>%
    drop_na() %>%
    ggplot(aes(year, life_exp, group = country)) +
    geom_line(aes(color = country), show.legend = TRUE, size = 0.5) +
    geom_point(aes(color = country), show.legend = TRUE, size = 1, shape = 16) +
    # geom_smooth(se = FALSE, span = 0.1) +
    scale_x_continuous(breaks = seq(1960, 2030, by = 10), expand = c(0, 1)) +
    coord_cartesian(clip = "off") +
    scale_y_continuous(limits = c(40, 70), expand = c(0, 0)) +
    scale_color_manual(breaks = c("Kenya", "Tanzania", "Uganda"), values = c(cb_palette)) +
    labs(
        title = "Life expectancy from 1960 to 2020",
        subtitle = "East Africa",
        x = "Year",
        y = "Average life expectancy (in years)",
        caption = "Data source: World Bank"
    ) +
    theme_classic() +
    theme(
        plot.title = element_text(margin = margin(b = 10)),
        plot.title.position = "plot",
        axis.text = element_text(margin = margin(b = 10)),
        plot.subtitle = element_text(size = 10, margin = margin(b = 15)),
        legend.position = c(0.9, 0.15), # Moves the legend to the top right inside chart area!
        legend.title = element_blank() # Removes the legend title, country
    )
# Save the gray variation

ggsave("figures/cb_east_africa_life_exp_1960-2020.png", width = 5.5, height = 4.5, unit = "in")
```

### Kenya versus China

```{r anim_kenya_china}
# Kenya vs. China
#
# Filter for Kenya & China
kenya_china_life_exp <- my_gapminder %>%
    filter(
        country %in% c("Kenya", "China")
    ) %>%
    drop_na()

# Kenya's vs. China's life expectancy plot
#
p_ke_china <- kenya_china_life_exp %>%
    ggplot(aes(year, life_exp, group = country, color = country)) +
    geom_line(show.legend = TRUE, linewidth = 1) +
    geom_point(fill = "white", shape = 21, size = 2, show.legend = TRUE) +
    # shape enables the white fill!
    scale_x_continuous(breaks = seq(1960, 2030, by = 10), expand = c(0, 1)) +
    coord_cartesian(clip = "off") +
    scale_y_continuous(limits = c(30, 80), expand = c(0, 0)) +
    labs(
        title = "Kenya's vs China's life expectancy from 1960 to 2020",
        x = "Year",
        y = "Average life expectancy (years)",
        caption = "Data source: World Bank"
    ) +
    theme_classic() +
    theme(
        plot.title = element_text(margin = margin(b = 10), color = "red", face = "bold"),
        # plot.title.position = "plot",
        axis.text = element_text(margin = margin(b = 10)),
        legend.title = element_blank(),
        legend.position = c(0.15, 0.95)
    ) +
    transition_reveal(along = year)

p_ke_china

ggsave("figures/kenya-china_life_exp_1960-2020.png", width = 5.5, height = 4.5, unit = "in")

animate(p_ke_china, width = 5.5, height = 4.5, unit = "in", res = 300)

anim_save("figures/kenya_v_china_lifeExp_line_animated_world_bank.gif")
```



# NORMALIZED

This section looks into the anomalies in chnages in life expectancy that Kenya has experienced between 1960 and 2020. The normalized_range is between 1980 and 1990!

```{r}
life_exp_diff <- my_gapminder %>%
    mutate(
        normalized_range = year >= 1980 &
            year <= 1990,
        norm_life_exp = sum(life_exp * normalized_range) / sum(normalized_range),
        life_exp_diff = life_exp - norm_life_exp
    ) %>%
    filter(country %in% c("Kenya", "China")) %>%
    select(country, year, life_exp_diff)

life_exp_diff %>%
    ggplot(aes(year, life_exp_diff, group = country)) +
    geom_line() +
    scale_x_continuous(breaks = seq(1960, 2030, by = 4)) +
    scale_y_continuous(limits = c(-34, 5))


ggsave("figures/kenya_normalized_life_exp.png")
```

## STATS ANALYSIS

H0: There is no statistically significant difference in life expectancy between Kenya and the rest of East Africa
H1: There is a statistically significant difference in life expectancy of Kenya compared to the rest of East Africa


```{r}
life_exp_diff %>%
    arrange(life_exp_diff) %>%
    filter(life_exp_diff < -10 | life_exp_diff > 5) %>%
    print(n = 33)
```


# COMMENTARY

-   In the code line `geom_point(fill = "", color = "", shape = 21)`, the entry `shape = 21` enables the plot **symbol = circle** and without it we do no get the white fill! Meeeh! This code line does not really apply in this plot it is borrowed by the idea is the same!
-   I dont have to include the two `geom_line` and `geom_point`. I only have `geom_point` for the sake of testing the various shapes!
-   Ignore my love (infatuation? Anyone?) for the drammatic `!`.
-   the drammatic difference between the charts is enable by the minute change in code line 149 above!! The beauty of R-programming, huh?axis.text
-   Finally, I have about 10 minutes to ride to town and catch the ARSENAL's #TeamKubwa revenge game! HAHAHA! BYE BYE FOR NOW WADAU!

# UPDATES

Game ya #TeamKubwa imeweza kuweza!!!!

<font color = "darkred">#COYG</font>.
No wonder, <font color = "darkred">LONDON</font> is <font color = "darkred">RED</font>!!

```{r}
kenya_china <- my_gapminder %>%
    filter(country %in% c("Rwanda", "Tanzania")) %>%
    drop_na()

t.test(data = kenya_china, life_exp ~ country)
```
