---
title: 'T-test and linear regression for '
author: 'William Komu'
date: '`r Sys.Date()`'
document:
  pdf_document:
    self_contained: false
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE,
    message = FALSE,
    warning = FALSE
)
```

```{r load_pkgs}
library(gapminder)
library(tidyverse)
library(ggthemes)
# attach(gapminder)
```

```{r load_color_palette}
cb_palette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
cbb_palette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

```{r summary_gp}
summary(gapminder)
```


```{r}
ke_sa <- gapminder %>%
    select(country, continent, year, lifeExp) %>%
    filter(country == "South Africa" |
        country == "Kenya")
```

### T-test
H0: There is no statistically significant difference between life expectancy in South Africa and Kenya
H1: There is a statistically significant difference between life expectancy in South Africa and Kenya

```{r, include = TRUE}
t.test(data = ke_sa, lifeExp ~ country)
```

### Results
There is no **<font color="blue"> statistically significant difference</font>** in mean life expectancy between Kenya and South Africa (p-value = 0.5667)!

### Change in life expectancy for a select countries

The comparision of the life expectancy for Kenya and the rest of the world is as contrasting as contrasting can be!!!

***<font color = 'red'>BOOOOM SHAKALAKA!!!!</font>***

```{r}
gapminder %>%
    select(year, country, lifeExp) %>%
    filter(country %in% c("Kenya", "India", "South Africa", "Singapore", "Ireland", "United States")) %>%
    ggplot(aes(year, lifeExp, group = country, color = country)) +
    geom_line(aes(color = country), linewidth = 1) +
    scale_color_manual(values = cbb_palette) +
    scale_x_continuous(limits = c(1950, 2010), expand = c(0, 0)) +
    coord_cartesian(clip = "off") +
    theme_classic() +
    theme(legend.position = "right") +
    labs(title = "Change in life expectancy from 1950 to 2007", x = "Year", y = "Life Expactancy (years)")

ggsave("figures/change_life_expectancy_frm_1950s-2007.png", width = 7, height = 6.5, units = "in", dpi = 300)
ggsave("figures/change_life_expectancy_frm_1950s-2007.pdf", width = 7, height = 6.5, units = "in", dpi = 300)
```

```{r}
gapminder %>%
    arrange(year)
```
