---
title: "Using ggplot2 to recreate a line plot of annual temperature anomalies-CC217"
author: 'William Komu'
date: '`r Sys.Date()`'
document: pdf_doc
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE,
    message = FALSE,
    warning = FALSE
)
```

```{r load_pkgs}
library(tidyverse)
library(ggplot2)
library(glue)
library(lubridate)
library(gganimate)
```

## Session Summary

"Pat recreates animated climate spirals of the annual temperature anomalies by month using NASA’s GISS data using tools from the ggplot2 and gganimate R packges. This figure shows the deviation in annual global mean temperatures from the normalized temperatures of 1951 to 1980 as a line plot. The lines are colored according to the year. He uses `transition_reveal`, `transition_manual`, `coord_polar`, `geom_rect`, `geom_line`, `scale_color_viridis_c`, and a smattering of arguments from the `theme` function to create this provocative visual. All of this is done in R with the help of RStudio."

### Load data and tidy

We need to load the data and tidy it to put the data in a long format... `pivot_longer`
The code `mutate(month = factor(month, levels = month.abb))` turns the months into numerical order rather than the alphabetical order they come out with in the first instance!


```{r load_data}
t_diff <- read_csv("/home/iamlnx/Documents/GitHub/nasoma_R/climate/data/GLB.Ts+dSST.csv", skip = 1, na = "***") %>%
    select(year = Year, month.abb) %>%
    pivot_longer(-year, names_to = "month", values_to = "t_diff") %>%
    mutate(month = factor(month, levels = month.abb))
```


### Preloaded vectors

The `month.abb` reports the abbreviated vec of the 12 months!

```{r test pre-loaded vec}
month.abb
```

### Programming `glue` objects

We can program some elements of the charts and insert them into our ggplot

```{r programming_vars}
this_year <- year(today()) - 1

min_year <- t_diff %>%
    summarise(year = min(year))

plot_tt <- glue("Global temparature change (1880-2022)")
```


### Rewind

This session is a continuation  of the previous two sessions CC217 - 218. It looks to make the spiral plot we made in session CC218 to go alive by animating it!

#### Changes and updates

Since we already have the spiral plot, albeit a static one, we will almost borrow the entire code from the previous session with minor changes! These are:
1. Remove the year 1879 from our `t_data` because it actully returns data for the 1880!
   `filter or select not 1879`
2. Change the ungroup by year and do it with ....
3. Use `mutate` to create a whole new column --- an index row_id using as follows:

`mutate(steps_ = 1: nrow(.))`

```{r engineering_last_Dec/next_Jan}
# last_Dec

# last_dec <- t_diff %>%
#     filter(month == "Dec") %>%
#     mutate(year = year + 1, month = "last_Dec")

# # next_Jan

next_jan <- t_diff %>%
    filter(month == "Jan") %>%
    mutate(year = year - 1, month = "next_Jan")

# Join/bind the three dfs
## Then pipe the joint_df into ggplot

t_data <- bind_rows(t_diff, next_jan) %>%
    mutate(
        month = factor(month, levels = c(month.abb, "next_Jan")),
        month_number = as.numeric(month),
        is_this_year = year == this_year
    ) %>%
    arrange(desc(month_number)) %>%
        filter(year != 1879) %>%
        mutate(step_number = 1:nrow(.))

t_data %>% tail()

annotation <- t_data %>%
    slice_max(year) %>% # Return to year because we got rid of the last_Dec
    slice_max(month_number)

# joint_t_diff %>%
#     filter(year == 2023)
```

### Initial draft plot

Let's plug the data to ggplot and display a skeleton of our draft plot, so that we can tell a kind of a story of the transition to the final draft.

Note: the `group` allows for plotting lines in `geom_line()`

### Final plot

#### Dummy tible
1. This tibble creates coordinates to pass into `geom_label` (_text too???) as labels to the outer red rings!!
2. The second is a df for months name and position

**NOTE:** Since ggplot2 takes any addition of `geom_()` as a layer, the sequence in which we put the `geom_()` argms determine the lower and uppeer, where the first arg is lower and the subseqs uppers. The dfs we create below will go on top of the `geom_line()`

```{r}
temp_lines <- tibble(
    x = 12,
    y = c(1.5, 2.0),
    labels = c("1.5\u00B0C", "2.0\u00B0C")
)

month_labels <- tibble(
    x = 1:12,
    labels = month.abb,
    y = 2.7
)
```

```{r spiral_anim_plt}
t_data %>%
    ggplot(aes(month_number, t_diff, group = year, color = year)) +
     geom_col(data = month_labels, aes(x = x + .5, y = 2.4), fill = "black", width = 1, inherit.aes = FALSE) +
     geom_col(data = month_labels, aes(x = x + .5, y = -2), fill = "black", width = 1, inherit.aes = FALSE) +
     geom_hline(yintercept = c(1.5, 2.0), color = "red") +
     geom_point(
         data = annotation, aes(
             x = month_number, y = t_diff,
             color = year
         ),
         size = 5,
         inherit.aes = FALSE
     ) +
      geom_label(
         data = temp_lines, aes(
             x = x, y = y,
             label = labels
         ),
         color = "red", fill = "black",
         label.size = 0,
         inherit.aes = FALSE
     ) +
     geom_text(
         data = month_labels,
         aes(
             x = x,
             y = y,
             label = labels
         ),
         inherit.aes = FALSE,
         color = "white",
         angle = seq(360 - 360 / 12, 0, length.out = 12),
     ) +
    geom_line() +
     geom_text(aes(x = 1, y = -1.3, label = "2022")) +
     scale_x_continuous(
         breaks = 1:12,
         labels = month.abb, expand = c(0, 0),
         sec.axis = dup_axis(name = NULL, labels = NULL)
     ) +
    scale_y_continuous(
        breaks = seq(-2, 2, 0.2),
        limits = c(-2, 2.7), expand = c(0, -0.7),
        sec.axis = dup_axis(name = NULL, labels = NULL)
    ) +
    scale_color_viridis_c(
        breaks = seq(1880, 2022, 20),
        guide = "none"
    ) +
    scale_size_manual(
        breaks = c(FALSE, TRUE),
        values = c(0.25, 1),
        guide = "none"
    ) +
    coord_polar(start = 2 * pi / 12) +
        # coord_cartesian(xlim = c(1, 12)) +
    labs(x = NULL, y = NULL) +
        # ggtitle(label = glue("Global temparature change since {min_year} by month")) +
    ggtitle(plot_tt) +
    theme(
            panel.background = element_rect(fill = "#444444", size = 1),
            plot.background = element_rect(fill = "#444444", color = "#444444"),
            panel.grid = element_blank(),
            axis.text.x = element_blank(),
            axis.text.y = element_blank(),
            axis.ticks = element_blank(),
            axis.title = element_text(color = "white", size = 13),
            plot.title = element_text(color = "white", hjust = 0.5, size = 15),
        ) +
        transition_reveal(along = step_number)

file_renderer(dir = ".", prefix = "climate_anima_plot", overwrite = FALSE)

# Save to help us view the effect of the alteration we do!

# ggsave("charts/climate_spiral_animated.png", width = 10, height = 5)
```


# Commentary

To keep my code chucks clean, I'll start doing commentaries - explanations of the ***not-so-obvious*** **code lines and tricks** or nifty tricks that I have a feeling are **outta this world** - at the end of the document.

### My adaptation:
In subsetting the data for the annotation trick, I adapted `is_this_year` because I am in 2023 which has no data other than the `last_Dec` we created!

In subsetting this year, I had to deduct one (-1) so that I can get last year's as this year! I will amend this line once i have a dataset that has data for 2023!
NOTE: last_Dec takes the previous years data and put it as a month 0, helping as to showcased a kind of transition/action in the edge of the plot!

### Read the end of df
Print the tail end of a df, where n = # of rows!
`data  %>% tail(n=#)`

```{r}
# t_diff %>% tail(n = 10)
```

```{r}
# ?theme(plot.title.position)
```


^[THIS IS A FOOTNOTE]

The format_document features does not always work when you have made a syntax error! Thus, it is a nice little **"bug buster"**, huh?
Change the angle of the text labels

`seq(0, 360 - 360 / 12, length.out = 12)`

# Get a circle

How about adding a circle behind our spiral to get the circular shape of the Hopkin's plot!



```{r}
# ggplot(data = month_labels, aes(x = x, y = 2.2), fill = "black") +
#     geom_col() +
#     coord_polar()
```

# CLEAN TO RUN WITH r

library(tidyverse)
library(ggplot2)
library(glue)
library(lubridate)
library(gganimate)

t_diff <- read_csv("/home/iamlnx/Documents/GitHub/nasoma_R/climate/data/GLB.Ts+dSST.csv", skip = 1, na = "***") %>%
    select(year = Year, month.abb) %>%
    pivot_longer(-year, names_to = "month", values_to = "t_diff") %>%
    mutate(month = factor(month, levels = month.abb))

# last_Dec

 last_dec <- t_diff %>%
     filter(month == "Dec") %>%
     mutate(year = year + 1, month = "last_Dec")

# # next_Jan

next_jan <- t_diff %>%
    filter(month == "Jan") %>%
    mutate(year = year - 1, month = "next_Jan")

# Join/bind the three dfs
## Then pipe the joint_df into ggplot

t_data <- bind_rows(t_diff, next_jan) %>%
    mutate(
        month = factor(month, levels = c(month.abb, "next_Jan")),
        month_number = as.numeric(month),
        is_this_year = year == this_year
    ) %>%
    arrange(desc(month_number)) %>%
    filter(year != 1879) %>%
    mutate(step_number = 1:nrow(.))

annotation <- t_data %>%
    slice_max(year) %>% # Return to year because we got rid of the last_Dec
    slice_max(month_number)

temp_lines <- tibble(
    x = 12,
    y = c(1.5, 2.0),
    labels = c("1.5\u00B0C", "2.0\u00B0C")
)

month_labels <- tibble(
    x = 1:12,
    labels = month.abb,
    y = 2.7
)

t_data %>%
    ggplot(aes(month_number, t_diff, group = year, color = year)) +
     geom_col(data = month_labels, aes(x = x + .5, y = 2.4), fill = "black", width = 1, inherit.aes = FALSE) +
     geom_col(data = month_labels, aes(x = x + .5, y = -2), fill = "black", width = 1, inherit.aes = FALSE) +
     geom_hline(yintercept = c(1.5, 2.0), color = "red") +
     geom_point(
         data = annotation, aes(
             x = month_number, y = t_diff,
             color = year
         ),
         size = 5,
         inherit.aes = FALSE
     ) +
      geom_label(
         data = temp_lines, aes(
             x = x, y = y,
             label = labels
         ),
         color = "red", fill = "black",
         label.size = 0,
         inherit.aes = FALSE
     ) +
     geom_text(
         data = month_labels,
         aes(
             x = x,
             y = y,
             label = labels
         ),
         inherit.aes = FALSE,
         color = "white",
         angle = seq(360 - 360 / 12, 0, length.out = 12),
     ) +
    geom_line() +
     geom_text(aes(x = 1, y = -1.3, label = "2022")) +
     scale_x_continuous(
         breaks = 1:12,
         labels = month.abb, expand = c(0, 0),
         sec.axis = dup_axis(name = NULL, labels = NULL)
     ) +
    scale_y_continuous(
        breaks = seq(-2, 2, 0.2),
        limits = c(-2, 2.7), expand = c(0, -0.7),
        sec.axis = dup_axis(name = NULL, labels = NULL)
    ) +
    scale_color_viridis_c(
        breaks = seq(1880, 2022, 20),
        guide = "none"
    ) +
    scale_size_manual(
        breaks = c(FALSE, TRUE),
        values = c(0.25, 1),
        guide = "none"
    ) +
    coord_polar(start = 2 * pi / 12) +
        # coord_cartesian(xlim = c(1, 12)) +
    labs(x = NULL, y = NULL) +
        # ggtitle(label = glue("Global temparature change since {min_year} by month")) +
    ggtitle(plot_tt) +
    theme(
            panel.background = element_rect(fill = "#444444", size = 1),
            plot.background = element_rect(fill = "#444444", color = "#444444"),
            panel.grid = element_blank(),
            axis.text.x = element_blank(),
            axis.text.y = element_blank(),
            axis.ticks = element_blank(),
            axis.title = element_text(color = "white", size = 13),
            plot.title = element_text(color = "white", hjust = 0.5, size = 15),
        ) +
        transition_reveal(along = step_number)
