---
title: 'Programming a lines plot in R to show climate change with and without animation (CC225)'
author: 'William Komu'
date: '`r Sys.Date()`'
document:
  pdf_document:
    self_contained: false
    toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE,
    message = FALSE,
    warning = FALSE
)

library(tidyverse)
library(gganimate)
options(browser = "firefox")
```

```{r}
month_anom <- read_table("/home/iamlnx/Documents/GitHub/nasoma_R/climate/data/merra2_seas_anom.txt", skip = 3) %>%
    select(month = Month, seas_anom) %>%
    mutate(
        month = as.numeric(month),
        month = base::month.abb[month]
    )

t_data <- read_csv("/home/iamlnx/Documents/GitHub/nasoma_R/climate/data/GLB.Ts+dSST.csv", skip = 1, na = "***") %>%
    select(year = Year, all_of(month.abb)) %>%
    pivot_longer(-year, names_to = "month", values_to = "t_diff") %>%
    inner_join(., month_anom, by = "month") %>%
    mutate(
        month = factor(month, levels = month.abb), month_anom = t_diff + seas_anom - 0.7
    ) %>%
    group_by(year) %>%
    mutate(ave = mean(month_anom)) %>%
    ungroup() %>%
    mutate(ave = if_else(year == 2022, abs(max(ave)), ave)) # %>%

# filter(year > 2010)

annotation <- t_data %>%
    slice_tail(n = 1)

current_year <- t_data %>%
    filter(year == 2022)
```
### Plot

```{r}
library(ggplot2)

p <- t_data %>%
    ggplot(aes(month, month_anom, group = year, color = ave)) +
    geom_line() +
    scale_color_gradient2(low = "darkblue", mid = "white", high = "darkred", midpoint = 0, guide = "none") +
    scale_y_continuous(breaks = seq(-3, 2, 1)) +
    scale_x_discrete(expand = c(0, 1)) +
    labs(
        x = NULL, y = NULL,
        title = "Temparature Anomaly (\u00B0 C)",
        subtitle = "(Difference between 1980 - 2015 annual mean)"
    ) +
    theme(
        panel.background = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.major.y = element_line(color = "gray", linetype = "dotted", size = 0.25),
        plot.title.position = "plot",
        plot.title = element_text(face = "bold"),
        plot.subtitle = element_text(color = "gray", size = 10),
        plot.margin = margin(t = 10, r = 15, b = 10, l = 10)
    )
# Save for real time tinkering:
p + geom_point(data = annotation, aes(x = month, y = month_anom), color = "#300606", size = 2.5) +
    geom_line(data = current_year, aes(month, month_anom), color = "darkred", linewidth = 1) +
    geom_text(data = annotation, aes(x = 11.5, y = -1.5), color = "darkred", label = "Dec 2022", hjust = 1)

ggsave("figures/c225_monthly_anomaly.png", width = 6, height = 4, units = "in")


# Animation
a <- p +
    geom_label(aes(x = 7, y = 0, label = year), fontface = "bold", label.size = 0) +
    transition_manual(year, cumulative = TRUE)

animate(a, width = 6, height = 4, units = "in", res = 300)
anim_save("figures/c225_monthly_anomaly.gif")
```

### Commentary

In the first attempt, we the plot gives us s strange outpput because the command to arrange months as fct has been overriden hence our months go back to being `char`

This is fixed by ensuring that the join occurs before the mutate for changing months!
This actually allows us to join the two mutate functions!

The issue with attempting to plot with  months as char is that the df is arranges the month alphabetically!!

### Adapt or die

I had to add geom_line for the `current_year` because my `year == 2022` is full year!
