---
title: "Mimic G-Trends Visuals for EPL Big-6 Searches in Kenya"
author: Willima Komu
date: "2022-02-14"
output: pdf_document
extra_dependencies:
      hyperref: ["unicode=true", "breaklinks=true"]
      lmodern: null
---

### Load packages
```{r, message=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)
library(readr)
library(emo)
library(tinytex)
library(readr)
library(ggthemes)
# library(here)

```

```{r}
cb_palette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
cbb_palette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")


```

### Load/import the dataset
```{r}
gunners_v_manutd <- read_csv("data/arsenal_v_man_u.csv", skip = 2, col_types = "Ddd") %>%
    clean_names() %>%
    rename(
        arsenal = arsenal_f_c_kenya,
        manutd = manchester_united_f_c_kenya,
        date = week
    )
head(gunners_v_manutd)
```
### Separate
We will need to seprate the Arsenal and Manchester's mentions into their respective data frames
#### Arsenal
```{r}
gunners <- gunners_v_manutd %>%
    select(date, arsenal) %>%
    mutate(club = "Arsenal") %>%
    rename(total_mention = arsenal)
head(gunners)
```
#### Man United
```{r}
manutd <- gunners_v_manutd %>%
    select(date, manutd) %>%
    mutate(club = "Man Utd") %>%
    rename(total_mention = manutd)
head(manutd)
```

### Append the **Arsenal** and **Man United** dfs
```{r}
gunners_v_manutd_df <- rbind(gunners, manutd)
head(gunners_v_manutd_df)
```

### Graph different data

We can plot the appended df

#### Line plot

The following code is the resulting line plot.


```{r}
guns_manutd_line_plot <- ggplot(data = gunners_v_manutd_df, aes(color = club, group = club)) +
    geom_line(aes(date, total_mention), size = 1) +
    scale_x_date(date_breaks = "1 year") +
    # geom_smooth(aes(date, total_mention, span = 0.01)) +
    theme_economist_white() +
    scale_color_manual(
        name = NULL,
        breaks = c("Arsenal", "Man Utd"),
        values = c("red", "purple")) +
    scale_x_date(date_labels = "%b %Y", date_breaks = "10 months")


guns_manutd_line_plot

ggsave("charts/google_trends/ref_date_manutd_arsenal_v_manutd_pop_kenya.png", plot = guns_manutd_line_plot, width = 7, height = 5)
```

`cbb_palette`
```{r}
guns_manutd_line_plot <- ggplot(data = gunners_v_manutd_df, aes(color = club, group = club)) +
    geom_line(aes(date, total_mention), size = 1) +
    scale_x_date(date_breaks = "1 year") +
    # geom_smooth(aes(date, total_mention, span = 0.01)) +
    theme_economist_white() +
    scale_color_manual(
        name = NULL,
        breaks = c("Arsenal", "Man Utd"),
        values = c(cb_palette)
    ) +
    scale_x_date(date_labels = "%b %Y", date_breaks = "10 months")


guns_manutd_line_plot

ggsave("charts/google_trends/cb_manutd_arsenal_v_manutd_ke.png", plot = guns_manutd_line_plot, width = 7, height = 5)

```

`cb_palette`
```{r}
guns_manutd_line_plot <- ggplot(data = gunners_v_manutd_df, aes(color = club, group = club)) +
    geom_line(aes(date, total_mention), size = 1) +
    scale_x_date(date_breaks = "1 year") +
    # geom_smooth(aes(date, total_mention, span = 0.01)) +
    theme_fivethirtyeight() +
    scale_color_manual(
        name = NULL,
        breaks = c("Arsenal", "Man Utd"),
        values = c(cbb_palette)) +
    scale_x_date(date_labels = "%b %Y", date_breaks = "8 months")


guns_manutd_line_plot

ggsave("charts/google_trends/cbb_manutd_arsenal_v_manutd_ke.png", plot = guns_manutd_line_plot, width = 7, height = 5)

```








### FIXING THE DATE

The read_csv enhances loading data, as one can denote the col_types. This workaround is key if you have to work with data at any point!









## How About Cumulative


```{r}
cum_manutd <- manutd %>%
    mutate(year = year(date), month = month(date)) %>%
    group_by(year) %>%
    mutate(cum_mention = cumsum(total_mention)) %>%
    ungroup()

#arsenal

cum_arsenal <- gunners %>%
    mutate(year = year(date), month = month(date)) %>%
    group_by(year) %>%
    mutate(cum_mention = cumsum(total_mention)) %>%
    ungroup()

arsenal_manutd_cum_df <- rbind(cum_arsenal, cum_manutd)

arsenal_manutd_cum_df %>%
    ggplot(aes(year, cum_mention, group = year, color =club)) +
    geom_point()

```










#### Side note on saving the plots

We need to save it to the charts folder so that it can display write here on vsc, which makes it easier to edit various elements and get a real-time output! This allows us to test various changes in a more efficient manner. Normally, the plot is opened in  a new window which is tedious to go view. Although you can set vsc to display viz side-by-side to the wspace, I have found the required settings quite ellusive to retrace! Thus, this is a trick I stumbled upon the other day!
However, you may not need it if you have enabled preview, but the preview feature is a nuisance that you may not need especially when running code!













































MY BATTLE WITH EMOJIS AND RMARKDOWN
#### Test github_document output renders emojis
<!--
```{r}
emo::ji("star")
```

:smile_cat: -->

This does not fix the issue with rendering emojis on PDF!

# ```{r}
# wink_emoji <- "\U0001f609"
# utf8::utf8_print(wink_emoji)
# #> [1] "😉​"
# utf8ToInt(wink_emoji)
# #> [1] 128521
# ```

***SO FAR THE SOLUTION IS TO KNIT IT AS HTML THEN PRINT***

This solution kills hyperlinks or so I found out!
